### 2022-06-22 old-papers/draft3/old-models.tex
Added to paper and converted to using many samples rather than an average density distribution

\documentclass[fleqn,usenatbib]{mnras}

\usepackage[T1]{fontenc}

\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{colortbl}
\usepackage{dsfont}
\usepackage{siunitx}
\usepackage{bm}

% For the sideways figure
\usepackage{pdflscape}
\usepackage{afterpage}

\usetikzlibrary{arrows.meta}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
\newcommand{\jtd}[1]{ {\bf{\color{red} JTD: #1}} }
\newcommand{\unit}[1]{\bm{\hat{#1}}}

\newcommand{\parens}[1]{\left( #1 \right)}
\newcommand{\brackets}[1]{\left[ #1 \right]}
\newcommand{\quat}[1]{\widetilde{\bm{#1}}}

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

\title[Flyby Constraints on Asteroids Interiors]{Constraining the Interiors of Asteroids Through Close Encounters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Jack T. Dinsmore, Julien de Wit]{
Jack T. Dinsmore,$^{1}$\thanks{E-mail: jtdinsmo@mit.edu}
Julien de Wit$^{2}$
\\
% List of institutions
$^{1}$Department of Physics, Massachusetts Institute of Technology\\
$^{2}$Department of Earth, Atmospheric, and Planetary Science, Massachusetts Institute of Technology
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
  Knowledge of the interior distribution of an asteroid can reveal its composition and constrain its evolutionary history. However, most asteroid observational techniques are not sensitive to interior structure. We investigate the degree to which interior density distributions can be extracted from angular velocity data gathered during a close encounter. We derive the equations of motion for a rigid asteroid's orientation and angular velocity to arbitrary order and use them to generate synthetic angular velocity data for a representative asteroid on a close Earth encounter. Using Markov Chain Monte Carlo fits, we re-extract the density moments of the asteroid in a wide range of scenarios to measure the degree to which best-fitting parameter precision is affected. Specifically, we use many injection-retrieval tests to study fit precision's dependence on the asteroid's moment of inertia, observational precision and cadence, orbital parameters, and initial spin pole direction as well as the planet being encountered, the quantity of near-pericentre data, and the central body oblateness. Finally, we discuss the degeneracy between the density moments and the actual density distribution and propose three models to construct a representative density distribution from fit results.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
  minor planets, asteroids: general -- methods: data analysis
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) or the Las Cumbres Observatory (LCO). Light curve analysis has been performed on many asteroids to study their surface properties, but observations that are sensitive to the interior asteroid structure are rarer.

However, tidal torque acting on an asteroid and perturbing its angular velocity during an encounter presents an opportunity to constrain the interior mass distribution. Such angular velocity perturbations have been observed and used to extract asteroid properties in several cases, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and asteroid binaries (3905) Doppler and (617) Patroclus respectively. Altered rotational states altered by Solar tidal torque have been excluded for 1I/2017 U1 `Oumuamua \cite{KWIECINSKI2018170}.

The upcoming 2029 encounter of 99942 Apophis is of particular interest and much theoretical work has been devoted to accurate estimates of its orbital and physical properties and predictions of how its angular velocity will be affected by tidal torque \cite{yu2014numerical, hirabayashi2021finite,valvano2022apophis, Lee2022Apophis, souchay2014rotational, souchay2018changes}. This research builds on a larger body of theoretical work. The gravitational two-body system has been studied in the context of tidal torque to different orders and with several different methods \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. Further studies showed that the tidal torque, observed through angular velocity perturbations, is sensitive to asteroid interior density distribution \cite{Naidu_2015, Makarov2022ChaosOO, RICHARDSON199847, scheeres2004evolution}. However, density distribution features beyond the moment of inertia ratios have not yet been extracted for any asteroid encounters. More research is needed to study in what cases these effects are observable, and what factors generally inhibit observation of these new features.

In this paper, we address this question by deriving a new equation of motion for an asteroid encounter system which is quick to compute and valid to arbitrary order. In section \ref{sec:methods}, we describe a simulation used to integrate the equations of motion and produce synthetic data of angular velocity over time, followed by a Markov Chain Monte Carlo (MCMC) fit process which  extracts physical characteristics of the asteroid (specifically, density moments) from the fit data. We then describe three methods to generate full density distributions from the density moments. Section \ref{sec:results} demonstrates the efficacy of this spin-data-to-density-distribution pipeline and explores its key properties. Finally, in section \ref{sec:discussion}, we measure the uncertainties of the best-fitting parameter estimates as a function of various encounter properties to determine which properties most inhibit accurate extraction of density distributions. We also test the density distribution extraction methods on several sample asteroids.

\section{Methods}
\label{sec:methods}

Only some aspects of an asteroid's density distribution affect tidal torque interactions. These are the asteroid ``length'' 
\begin{equation}
  a_\mathcal{A}^2 = \frac{1}{\mu_\mathcal{A}} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) r^2
  \label{eqn:am}
\end{equation}
and (complex) ``density moments''
\begin{equation}
  K_{\ell m} = \frac{1}{\mu_\mathcal{A} a_\mathcal{A}^\ell} \int_\mathcal{A} d^3 r \rho_\mathcal{A}(\bm r) R_{\ell m}(\bm r),
  \label{eqn:klm}
\end{equation}
both of which are constant and only need to be computed once to generate data. Here, $\rho_\mathcal{A}(\bm r)$ is the asteroid density distribution, $\mu_\mathcal{A}$ is the asteroid mass, and $R_{\ell m}$ are the regular solid spherical harmonics (see appendix \ref{app:eom} for details). They produce a tidal torque of 
\begin{equation}
  \begin{split}
  \bm \tau = & G\frac{\mu_\mathcal{A}\mu_\mathcal{B}}{2}\left[\sum_{\ell m} a_\mathcal{B}^\ell J_{\ell m} \sum_{\ell' m'}a_\mathcal{A}^{\ell'}S^*_{\ell+\ell', m + m'} (\bm D) (-1)^{\ell'}\right.\\
  \times & \left.\sum_{m''=-\ell'}^{\ell'} \sqrt{\frac{(\ell'-m'')!(\ell'+m'')!}{(\ell'-m')!(\ell'+m')!}}  \mathcal{D}^{\ell'}_{m'm''}(\alpha, \beta, \gamma)^* \right. \\
  \times & \Big((i\unit x - \unit y)(\ell'-m''+1)K_{\ell',m''-1} \\
  & +(i\unit x+\unit y)(\ell'+m''+1)K_{\ell',m''+1}+2im''\unit z K_{\ell'm''}\Big) \Bigg],
  \end{split}
  \label{eqn:tidal-torque}
\end{equation}
where $\bm D$ is the position of the asteroid; $\alpha$, $\beta$, and $\gamma$ are $z-y-z$ Euler angles expressing the orientation of the asteroid; $\mu_\mathcal{B}$, $a_\mathcal{A}$, and $J_{\ell m}$ are the mass, length, and density moments of the central body respectively; and $S_{\ell m}$ are the irregular solid spherical harmonics. Equation \ref{eqn:tidal-torque} is derived in appendix \ref{app:eom}, assuming a rigid asteroid and no third-body perturbations. Throughout the paper, we refer to the ``inertial frame'' (the frame in which the orbit is fixed) and the ``body-fixed frame'' (the frame in which the asteroid is fixed), which are also defined in appendix \ref{app:eom}.

In section \ref{sec:sim}, we describe the simulation used to integrate the equations of motion. In section \ref{sec:uncertainty}, we define the uncertainty model used to add noise to the data, and we describe how posterior probability distributions (PPDs) for density moments were extracted from spin-pole data in section \ref{sec:fit}. Finally, we outline a general model to extract density distributions from density moments in section \ref{sec:general-density} and define specific model implementations in section \ref{sec:specific-density}.

\subsection{Simulation design}
\label{sec:sim}

We built a publicly accessible, custom simulation in C++ to produce angular velocity data as a function of time. This simulation requires as initial data (1) the orbital parameters of the asteroid $r_p$ (perigee distance) and $v_\infty$ (hyperbolic excess velocity); (2) the cadence of angular velocity observation $\Delta t$; (3) the central body moments $J_{\ell m}$, mass $\mu_\mathcal{B}$, and radius $a_\mathcal{B}$; (4) the initial asteroid angular velocity in the inertial frame $\bm \Omega_0$; (5) the asteroid length $a_\mathcal{A}$; and (6) the asteroid's density moments $K_{\ell m}$ and initial Euler angle $\gamma_0$. All parameters except (6) are assumed to be known to high accuracy. One can imagine that $a_\mathcal{A}$ is approximated by light-curve analysis, but if not, it is still necessary to fix $a_\mathcal{A}$ or else the values of $K_{\ell m}$ are degenerate with $a_\mathcal{A}$.

We further assume that the asteroid is initially not tumbling. Thus, the angular velocity is aligned with a principal axis (assumed to be $\unit z$, which maximizes moment of inertia). This sets $\beta = 0$ and we can further choose $\alpha = 0$. Thus, only the Euler angle $\gamma_0$ is necessary to provide initial data for the simulation.

We begin our simulation at $D = 10 r_p$, with velocity predicted by Kepler's laws given by the orbital parameters. Since the leading order of the equations of motion is $\ell' = 2, \ell = 0$, this corresponds roughly to a torque of $10^{-3}$ times the maximum torque at perigee. Unless otherwise indicated, the simulation is terminated at $D=10 r_p$ as well.

With the simulation inputs specified, the equations of motion are integrated via the Runge-Kutta fourth order method, with a variable time step
\begin{equation}
  dt = dt_\text{min} + 10^{-3}(dt_\text{max} - dt_\text{min}) \brackets{\parens{\frac{D}{r_p}}^3 - 1}.
\end{equation}
The parameters $dt_\text{max}$ and $dt_\text{min}$ (20 and 10 seconds respectively) were chosen such that the numerical integration error was $\sim$100 times the floating point error, and that neighbouring values of $K_{\ell m}$ yielded significantly different spin pole data compared to floating point error. The data used to choose this $dt$ was obtained using the reference asteroid configuration, described in appendix \ref{app:reference-configs}.

\subsection{Uncertainty model}
\label{sec:uncertainty}

To add noise to data generated via the above simulation, we use the following uncertainty model. Each asteroid spin vector $\bm \Omega$ is assumed to be uncorrelated with other spin vectors, and we model uncertainty in the orientation and in the period as also uncorrelated. Consider a true spin vector $\bm \Omega^*$. For the sake of description, we work in coordinates in which $\bm \Omega^* \parallel \unit z$. Then, expressing the observed spin vector $\bm \Omega$ in spherical coordinates, we draw the polar angle from a normal distribution with standard deviation $\sigma_\theta$ centred on zero and the azimuthal angle from a uniform distribution. We also draw the ratio $\rho=\Omega/\Omega^*$ from a log-normal distribution centred on one, with width $\sigma_\rho$. This width is related to the uncertainty in asteroid period $\sigma_P$ by $\sigma_P=P \sigma_\rho$ for small $\sigma_\rho$. Explicitly, the probability density function (PDF) of $\rho$ is 
\begin{equation}
  P(\rho) = \frac{1}{\rho\sqrt{2\pi \sigma_\rho^2}} \exp\parens{-\frac{\ln^2\rho}{2\sigma_\rho^2}}.
\end{equation}
See figure \ref{fig:uncertainty-model} for an illustration of the uncertainty model.

\begin{figure}
  \centering
  \begin{tikzpicture}
  \draw[-{Latex[length=3mm]}] (0, 0) -- (-4, 0) node[anchor=east] {$\unit X$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (1, -1.5) node[anchor=west] {$\unit Y$};
  \draw[-{Latex[length=3mm]}] (0, 0) -- (0, 4) node[anchor=south] {$\unit Z$};

  \draw[line width=0.5mm, -{Latex[length=3mm]}] (0, 0) -- (-3, 3) node[anchor=south] {$\bm \Omega^*$};
  \draw[dashed] (-3, -2) -- (-3, 3);
  \draw[dashed] (-3, -2) -- (0, 0);

  \draw[-{Latex[length=3mm]}] (0, 0) -- (-1.25, 3.35) node[anchor=south west] {$\bm \Omega$};
  \draw[rotate around={-50:(-3,3)}] (-3,3) ellipse (1.4 and 2);
  \draw (-3, 3) arc (110:93:6);
  \draw (-2, 3.5) node[anchor=center] {$\theta$};
  \draw (-0.5, 2.5) node[anchor=center] {$\rho \Omega^*$};

  \end{tikzpicture}
  \caption{Diagram in the inertial frame of the uncertainty model used to define the probability that the true spin vector $\bm \Omega^*$ should be observed as $\bm \Omega$. The parameter $\theta$ is drawn from a Gaussian with width $\sigma_\theta$, and $\rho$ is drawn from a log normal distribution with width $\sigma_\rho$.}
  \label{fig:uncertainty-model}
\end{figure}

The log likelihood resulting from this uncertainty model is (excluding additive constants)
\begin{equation}
  \begin{split}
  \ln \mathcal{L} = -\frac{1}{2}\sum_{i = 0}\Bigg[&\frac{\cos^{-1} (\bm \Omega_i^* \cdot \bm \Omega_i/(\Omega_i^* \Omega_i))^2}{\sigma_\theta^2}\\
  &+\frac{\ln \parens{\Omega_i /\Omega_i^*}^2}{\sigma_\rho^2} + 2\ln\frac{\Omega_i}{\Omega_i^*}\Bigg].
  \end{split}
  \label{eqn:log-likelihood}
\end{equation}
where $\Omega_i$ is the $i$th spin vector in the data set.

This model was chosen because it separates spin pole and period uncertainty. Therefore, if one is more precisely determined by measurement, $\sigma_\theta$ and $\sigma_\rho$ can be adjusted separately in accordance.

\subsection{Extracting density moments from spin data}
\label{sec:fit}
Given synthetic data, an Affine Invariant MCMC Ensemble sampler was used to generate PPDs from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), and were bounded by $|\gamma_0| < \pi/4$, and bounds on $K_{2 m}$ given in equation \ref{eqn:parameter-bounds}. Note that $\gamma_0$ is degenerate with $\gamma_0 + \pi/2$ because both of these align the asteroid principal axes with the body-fixed axes. The other bounds were $|K_{3m}| < 1$. In general, spin data is most sensitive to $\gamma_0$ and $K_{2m}$, which we call the ``first-order parameters.'' We call $K_{3m}$ the ``second-order parameters.''

The MCMC was determined to converge when the fractional change in autocorrelation time (computed every 100 iterations) was one percent, and the number of iterations computed so far was more than 100 times the autocorrelation time. The MCMC fit also was set to terminate if more than $10^5$ iterations were run, but this only occurred for fits in which the data was too low-quality to resolve the parameters. $\sim 10^4$ iterations was often sufficient, which generally consumed about 7 hours of computation time on a super computer running 16 threads on 8 cores.

Before the MCMC was run, local minima in the likelihood were found via the Nelder-Mead algorithm implemented in \texttt{scipy} \cite{Gao2012}. Generally, only one local minimum existed, except when $K_{22}=0$ in which case rotational symmetry caused multiple values of $\gamma_0$ to be degenerate. Ensemble walkers were initialized near this local minimum, distributed by a Gaussian approximation of the likelihood, as determined via the Hessian of the likelihood at the minimum. Due to the high sensitivity of the angular velocity data to density moments, the minimization procedure sometimes failed to isolate the minimum likelihood. Therefore, a simpler simulation without the $K_{3m}$ terms of equation \ref{eqn:tidal-torque} was first used to minimize likelihood as a function of the first-order parameters $\gamma_0$ and $K_{2m}$, and then the full simulation was used to find the second-order parameters $K_{3m}$ with the first-order parameters fixed.

To further ensure convergence, we minimized with respect to data truncated after perigee at double the perigee distance. This cut-off was manually chosen. The minimum was then further refined by minimizing based on the full data, with the previous minimum as the initial estimate.

\subsection{Extracting density distributions from moments: general model design}
\label{sec:general-density}

The asteroid density distribution $\rho_\mathcal{A}(\bm r)$ is not uniquely constrained via tidal torque interactions because only the density moments $K_{\ell m}$ contribute to equation \ref{eqn:tidal-torque}. However, by making sufficient assumptions about the density distribution, we can nevertheless estimate $\rho(\bm r)$ from $K_{\ell m}$. We outline three different assumptions which constitute three models for extracting density distributions from density moments. All are (partially) linear to facilitate computational ease and uncertainty propagation. One of them can be chosen, or another created, depending on what features are desired to be represented in the density distribution.

All three of these models assume that $a_\mathcal{A}$ and the asteroid shape is known (for example, by light curve or radar data). Uncertainty on this shape estimate is assumed to be small. The models then produce a density distribution $\rho_\mathcal{A}(\bm r)$ with uncertainty from $K_{\ell m}$ propagated to uncertainty in $\rho_\mathcal{A}$. If the asteroid shape is not known, a fourth model defined in appendix \ref{app:find-surface} can be used to extract it assuming uniform density. In section \ref{sec:results}, we were forced to fix $a_\mathcal{A}$ rather than fit for it because $a_\mathcal{A}$ is degenerate with scaling $K_{3m}$. However, $a_\mathcal{A}$ contains information about the density distribution as shown by its definition in equation \ref{eqn:am}. It might therefore be unreasonable to assume that it is exactly determined by light-curve or radar analysis. If $a_\mathcal{A}$ is too uncertain to be used as a known quantity when extracting density distributions from the moments, then the following models are flexible enough that the $a_\mathcal{A}$ constraint can be removed.

@@

\subsection{Extracting density distributions from moments: specific models}
\label{sec:specific-density}

@@

We now explore a model that seeks to restrict the space of allowed density distributions a different way: we allow only the density distributions with $\nabla^2 \rho = 0$ inside the asteroid (the harmonic distributions). We  call this  the ``harmonic'' model. Any harmonic function can be expanded as 
\begin{equation}
  \rho(\bm r) = \sum_{\ell m} \Theta_{\ell m}\frac{R_{\ell m}^*(\bm r)}{a_\mathcal{A}^\ell} 
  \label{eqn:harmonic-rho}
\end{equation}
where the terms which lead to $\rho \rightarrow \infty$ at the origin have been removed and $\Theta_{\ell m}$ are free (complex) parameters. Since $\rho$ is real, we have $\Theta_{\ell m}=(-1)^m \Theta_{\ell,-m}^*$. By setting a maximum on $\ell$, we restrict the number of degrees of freedom to $(\ell_\mathrm{max}+ 1)^2$. Choosing the same maximum $\ell$ as the maximum $\ell$ for the $K_{\ell m}$ moments, we have $m=n-1$ coefficients $\Theta_{\ell m}$ which can be stacked into an $m$-dimensional vector $\bm \Theta$.

@@

The above two simple models produce smooth density distributions. In this section, we define a more complicated model which identifies discrete regions differing from the overall density. We call this model the ``lumpy'' model.

@@

We use a non-linear process to choose $\bm x_i$ and $\mu_i$ so that $K_{\ell m}^i$ and $a_i$ can then be extracted via the linear process described in section \ref{sec:general-density}. We must incorporate the centre-of-mass and total-mass constraints
\begin{equation}
  \mu_0 \bm x_0 + \sum_{i=1}^N \mu_i \bm x_i = 0, \qquad \mu_0 + \sum_{i=1}^N \mu_i = \mu_\mathcal{A}.
  \label{eqn:lump-constraints}
\end{equation}
There are also additional constraints, such as that $\bm x_i$ should lie inside the asteroid, which can be enforced manually. Combining both constraints in equation \ref{eqn:lump-constraints}, we can eliminate $\mu_0$ and vary only $\mu_i$ and $\bm x_i$ for $i \geq 1$.

@@

\section{Results}
\label{sec:results}

\subsection{Density moment extraction}
\label{sec:example-fit}
To demonstrate our fit process, we present example results of the MCMC fit described in section \ref{sec:fit}, applied to encounter data of the asymmetric reference asteroid. The fit results were consistent with the true density moments, and produced consistent data.

\begin{figure*}
  \centering
  \includegraphics[width=0.6\textwidth]{figs/example-residuals.pdf}
  \caption{Data, best-fitting results, and residuals for a fit to synthetic data simulated for an asymmetric reference asteroid. Uncertainty bands are also shown.}
  \label{fig:example-residuals}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.85\textwidth]{figs/example-corner.png}
  \caption{PPDs extracted from synthetic encounter data for the asymmetric reference asteroid. Samples from the MCMC fit are shown as individual points, and the contours enclose 1, 2, and 3$\sigma$ confidence regions. True values are shown as blue lines.}
  \label{fig:example-corner}
\end{figure*}

Figure \ref{fig:example-residuals} shows the spin data with noise added. The best-fitting model is overlaid in the top pane and residuals are shown the bottom pane. Specifically, we show the model whose parameters are the means of the PPDs. Uncertainties are plotted on the residuals corresponding to the square root of the diagonal entries of the covariance matrix (correlations not included). The fit results are clearly consistent with the data.

Figure \ref{fig:example-corner} shows a corner plot of the ten fitted PPDs, marginalized to functions of one (histograms) or two (contours) variables. The true parameters are shown as lines. Note that the true parameters usually lie within 1 or 2$\sigma$ of the $\Delta K_{\ell m} = 0$, where $\Delta K_{\ell m}$ is the difference between the posterior $K_{\ell m}$ and the true $K_{\ell m}$. The PPDs are generally Gaussian and sometimes show correlation between parameters, but no continuous degeneracy occurs. We performed 48 independent minimizations of the likelihood before the MCMC fit began, each with an initial point chosen randomly in the parameter space. All converged to the same minimum, demonstrating that the model lacks discrete degeneracy as well.

\subsection{Density distribution extraction}
\label{sec:asym-density}
To demonstrate our three density distribution extraction models, we show density distributions and uncertainties extracted for the asymmetric reference asteroid in figure \ref{fig:uniform-density-results}. These distributions were made given the density moment PPDs extracted above, with $a_\mathcal{A}$ fixed. The surface is fixed as a triaxial ellipsoid with semi-axis lengths such that the true values of $K_{\ell m}$ are consistent with uniform density (equation \ref{eqn:ellipsoid-axes}).

\begin{figure*}
  \centering
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-lumpy.pdf}

  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-likelihood.pdf}\hfill
  \includegraphics[width=0.3\textwidth]{figs/asym-ell-unc-harmonic.pdf}\hfill\hspace{0.3\textwidth}

  \caption{Density distributions (\textit{top}) and uncertainties (\textit{bottom}) extracted by the likelihood (\textit{left}), harmonic (\textit{middle}), and lumpy with $d=N=1$ (\textit{right}) models from data generated for a uniform-density, asymmetric reference asteroid. Density is normalized so that mean density is one.}
  \label{fig:uniform-density-results}
\end{figure*}

Also in each pane of figure \ref{fig:uniform-density-results} is $\Delta K_{\ell m}$, the squared magnitude of difference between the fitted density moments and the density moments of the final distribution. This difference is non-zero due to numerical error in the likelihood model case, which is guaranteed to exactly reproduce $K_{\ell m}$. The harmonic and lumpy models have no such guarantee, however.

The likelihood and harmonic models produce nearly identical density distributions. This is generally true for uniform density asteroids, which have harmonic density distributions. Consequentially, $\Delta K_{\ell m}$ is low for the harmonic model. This similarity lends confidence to both models, because if one were unexpectedly biased towards certain distributions, the other would have to be equally biased in the same way in this case to produce the same distribution, which appears unlikely given that the models share little in common.

The distributions are non-uniform for the harmonic and likelihood models (driven by error in the estimates of $K_{3m}$), but the non-uniformity is limited. Note that most of the asteroid has density close to the mean density (one) in both cases, with densities of 40\% or 160\% of the mean attained only in small regions. The lumpy model yields a much more uniform distribution because it is designed to produce uniform lumps. However, it does not produce a density distribution consistent with the moments extracted from spin data (i.e., $\Delta K_{\ell m}$ is large). This is because the lump found extends outside the asteroid, violating the assumptions of the model. This is a sign that the lump found does not really exist inside the asteroid.

The size of uncertainty is such that the density distribution is consistent with uniform for the harmonic and likelihood models (within 1.5$\sigma$). Uncertainty is generally larger towards the edges of the distribution, both for these shapes and for the shapes shown later.

\section{Discussion}
\label{sec:discussion}

\subsection{Spin evolution}
\label{sec:spin-evolution}
In figure \ref{fig:example-data}, we present example spin data generated via our simulation. A population of one thousand asteroids with identical initial conditions except for $\gamma_0$, $K_{20}$, and $K_{22}$ were simulated on a close Earth encounter. The exact parameters used were the symmetric and asymmetric cases described in appendix \ref{app:reference-configs}. Bands containing 68.3\%, 95.5\%, and 99.7\% of the population's spin are shown, as is the spin of the reference asteroids in black.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/nominal-data-sym.pdf}
  \includegraphics[width=\linewidth]{figs/nominal-data-asym.pdf}
  \caption{Angular velocity data simulated for the symmetric (above) and asymmetric (below) reference asteroids. The true angular velocity evolution is shown as a black line. Also plotted is the deviation of the data for PPD-distributed perturbations to the asteroid shape (bands). Bands contain 68.3\%, 95.5\%, and 99.7\% of the 1000 simulations run.}
  \label{fig:example-data}
\end{figure}

The initial values of $\gamma_0$, $K_{20}$, and $K_{22}$ were distributed proportional to the PPDs extracted via the MCMC fit shown in section \ref{sec:example-fit}. The PPDs were widened by a factor of 1000 to make the band widths visible. Therefore, the scale of the bands in figure \ref{fig:example-data} have little meaning in an absolute sense, but they are meaningful when comparing two bands or two times in the same band.

The figure illustrates the sensitivity of spin data to asteroid density moments and $\gamma_0$; Before perigee, all asteroids had similar angular velocities, but after perigee the angular velocities of the population diverged. The spin data at perigee is informative and shows non-trivial behaviour. After perigee, the asteroid population tumbles according to torque-free precession with different oscillation periods. This dramatically increases the spread of angular velocities within the distribution and allows for more precise refinement of the density moments. The oscillation periods observable during the torque-free precession phase of the spin data (about 5 hours after perigee) can be used to constrain $K_{22}$ and $K_{20}$, while the phase is set by the asteroid dynamics near perigee and therefore constrains the higher-order terms.

\subsection{Dependence of parameter uncertainty on encounter properties}
\label{sec:fit-uncertainty}

In this section, we discuss how the precision of the best-fitting density moments is affected by various encounter properties. We call this precision ``posterior uncertainty,'' or $\sigma(K_{\ell m})$, defining 1$\sigma$  posterior uncertainty as the range containing 68.27\% of the PPD centred on the median. 2$\sigma$ posterior uncertainty is defined likewise for 95.45\% of the PPD. These are roughly equivalent to 1 or 2 times standard deviation of the PPD respectively. We use the configuration of the asymmetric reference asteroid (appendix \ref{app:reference-configs}) unless otherwise stated.

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all1.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of (left to right) perigee, excess velocity, spin pole uncertainty, and period uncertainty. The vertical dashed line indicates the reference asteroid values. The red vertical lines indicate when $\sigma(K_{3m}) =0.01$.}
    \label{fig:scan-perigee}
    \label{fig:scan-vex}
    \label{fig:scan-theta}
    \label{fig:scan-rho}
    \end{figure}
  \end{landscape}
}

\afterpage{
  \begin{landscape}
    \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figs/scan-all2.pdf}
    \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of (left to right) asteroid length, rotational period, observational cadence, and data gap at perigee. The vertical dashed line indicates the reference asteroid values. The red vertical lines indicate when $\sigma(K_{3m}) =0.01$.}
    \label{fig:scan-am}
    \label{fig:scan-cadence}
    \label{fig:scan-period}
    \label{fig:observation-gap}
    \end{figure}
  \end{landscape}
}

There is no clear threshold above which posterior uncertainty is ``too high.'' However, most asteroid distributions tested in this paper have $K_{3 m} \lesssim 0.01$. We therefore call $\sigma(K_{3 m}) \approx 0.01$ the ``posterior uncertainty threshold,'' beyond which $K_{3 m}$ are unresolved. This should be understood as a rule of thumb, not a precise statement.

>>>>

\subsection{Translating constraints on density moments into constraints on the density distributon}
\label{sec:density-distro}

The asteroid density distribution $\rho_\mathcal{A}(\bm r)$ is not uniquely constrained via tidal torque interactions because only the density moments $K_{\ell m}$ contribute to equation \ref{eqn:tidal-torque}. However, by making sufficient assumptions about the density distribution, we can nevertheless estimate $\rho(\bm r)$ from $K_{\ell m}$. We outline three different assumptions which constitute three models for extracting density distributions from density moments. 
Our goal in investigating different families of models to translate constraints on the density moments into constraints on the density distribution is to reveal the major features in the density distribution that are consistently present, independent of the family of mapping model used, despite the problem underdetermination (in a similar fashion to \cite{de2012towards}).

All are (partially) linear to facilitate computational ease and uncertainty propagation. All three of these models assume that $a_\mathcal{A}$ and the asteroid shape is known (for example, by light curve or radar data). Uncertainty on this shape estimate is assumed to be negligible compared to the density moment uncertainty and their contributions to to the density distribution. The models then produce a density distribution $\rho_\mathcal{A}(\bm r)$ with uncertainty from $K_{\ell m}$ propagated to uncertainty in $\rho_\mathcal{A}$. If the asteroid shape is not known, a fourth model defined in appendix \ref{app:find-surface} can be used to extract it assuming uniform density. In section \ref{sec:results}, we were forced to fix $a_\mathcal{A}$ rather than fit for it because $a_\mathcal{A}$ is degenerate with scaling $K_{3m}$. When extracting density distributions, multiple $a_\mathcal{A}$ values may be tested and the multiple produced density distributions can be compared to highlight common features.

@@

@@

We now explore a model that seeks to restrict the space of allowed density distributions in a different way: we allow only the density distributions with $\nabla^2 \rho = 0$ inside the asteroid (the harmonic distributions). We  call this  the ``harmonic'' model. Any harmonic function can be expanded as 
\begin{equation}
  \rho(\bm r) = \sum_{\ell m} \Theta_{\ell m}\frac{R_{\ell m}^*(\bm r)}{a_\mathcal{A}^\ell} 
  \label{eqn:harmonic-rho}
\end{equation}
where the terms which lead to $\rho \rightarrow \infty$ at the origin have been removed and $\Theta_{\ell m}$ are free (complex) parameters. Since $\rho$ is real, we have $\Theta_{\ell m}=(-1)^m \Theta_{\ell,-m}^*$. By setting a maximum on $\ell$, we restrict the number of degrees of freedom to $(\ell_\mathrm{max}+ 1)^2$. Choosing the same maximum $\ell$ as the maximum $\ell$ for the $K_{\ell m}$ moments, we have $m=n-1$ coefficients $\Theta_{\ell m}$ which can be stacked into an $m$-dimensional vector $\bm \Theta$.

@@

The above two simple models produce smooth density distributions. In this section, we define a model which identifies discrete regions differing from the overall density. We call this model the ``lumpy'' model.

@@

We use the following non-linear process to choose $\bm x_i$ and $\mu_i$ so that $K_{\ell m}^i$ and $a_i$ can then be extracted via the linear process described above. We must incorporate the centre-of-mass and total-mass constraints
\begin{equation}
  \mu_0 \bm x_0 + \sum_{i=1}^N \mu_i \bm x_i = 0, \qquad \mu_0 + \sum_{i=1}^N \mu_i = \mu_\mathcal{A}.
  \label{eqn:lump-constraints}
\end{equation}
There are also additional constraints, such as that $\bm x_i$ should lie inside the asteroid, which can be enforced manually. Combining both constraints in equation \ref{eqn:lump-constraints}, we can eliminate $\mu_0$ and vary only $\mu_i$ and $\bm x_i$ for $i \geq 1$.

@@
